CKY Parsing With Independence Constraints

### 2015-05-16 paper.tex
Copy draft LaTeX export to paper.tex

% Created 2015-02-28 Sat 15:32
\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{float}
\usepackage{acl2015}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fontspec}
\usepackage{xunicode}

\usepackage{forest}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\BigO}[1]{\ensuremath{\operatorname{O}\bigl(#1\bigr)}}
\date{\today}
\title{CKY Parsing With Independence Constraints}
\begin{document}

@@

The CKY algorithm is a \BigO{|G|n^3} dynamic programming
algorithm for finding all of the possible derivations of a sentence in
a context-free language. Its complexity depends on both the sentence
length $n$ and the size of the grammar $|G|$. Methods for improving
parse accuracy typically increase the size of the grammar 
\cite{Klein2003,Petrov2007} or even the exponent of $n$ \cite{Eisner1999}. 
More powerful “deep” grammar formalisms multiply the computational
complexity even more \cite{Bangalore1999}.

A common technique for speeding up such parsers is coarse-to-fine
parsing, where input is first parsed using a much simpler (and thus
smaller) grammar, and the content of the chart is then used to
constrain the search over the final grammar
\cite{Torisawa2000,Charniak2005,Petrov2007}. Even with a much smaller
grammar, the CKY algorithm may be expensive—Roark et al. (2012)
report that the initial CKY step in the Berkeley Parser takes half of
the total parse time.

Other techniques can be used to prune cells in the chart. Roark et al.
(2012) use a finite-state model to label words that do/don’t begin/end
spans, and skip cells that don’t satisfy the labels. Bodenstab et al.
(2011) directly apply a classifier to each cell to decide how many
spans to keep. Both approaches reduce the work done by the parser
while preserving accuracy.

We propose a novel type of top-down constraint for a CFG parser that
we call independence constraints, described in Section 2. In Section 3
we show how the CKY algorithm can be easily modified to accomodate
these constraints, and in Section 4 we describe a classifier which can
provide the constraints to a parser. We integrate the constraints into
the Stanford Parser CKY implementation and show the results in Section 5.

@@

\begin{figure}
\begin{forest}
  [S
   [NP [DT [ $_0$ This $_1$]]]
   [VP
    [VB [is $_2$]]
    [NP [DT [an $_3$]]
        [NN [example $_4$]]]]
   [{.} [{.} $_5$]]
  ]
\end{forest}
\caption{In this tree ‘This’ and ‘is’ are independent, while ‘is’ and ‘an’ are not.}
\label{fig:independence}
\end{figure}

@@

An example is shown in Figure \ref{fig:independence}. Here, ‘This’ and
‘is’ are independent, as are ‘example’ and ‘.’. The independent spans
are (‘This’), (‘is’, ‘an’, ‘example’), and (‘.’), with boundaries 1
and 4. The independent spans and independent span boundaries can be
derived straightforwardly from the definition of independent words:
the locations between consecutive words which are independent are the
independent span boundaries, and the independent spans are simply the
spans in between consecutive boundaries.

@@

The first complication is that if we assume that the independent boundaries
will be identified automatically, we must allow for errors. If a
location which is not an independent span boundary is given as one,
the parser will make an error it would not have otherwise. On the
other hand, if a location which is an independent span boundary is not
marked as such, the parser may account for this at the cost of not
achieving the minimum computation possible. By allowing for this
second type of error, the algorithm is made more robust, and allows
the independent boundary identification step to prioritize precision
over recall to lessen negative impact on the parser’s accuracy.

The second issue is caused by the binarization of the context-free
grammar used in the CKY algorithm. Because the CKY algorithm requires
a binary grammar, any rules in the original grammar that have more
than two symbols on the right-hand side must be converted into a
sequence of binary rules. The extra rules created in this process are
called intermediate or synthetic rules. The topmost span in particular
will usually need to be constructed in several steps; however, the
bottom-up nature of the algorithm means that it is impossible to
determine whether a given synthetic rule at any position in the CKY
chart will eventually be used to construct a complete span at the top
of the tree (for arbitrary grammars—if the grammar is left- or
right-binarized, only cells on the left or right edge of the chart can
affect the top span).

The combination of these two issues means that in order to correctly
parse a sentence when an independent span boundary is missing from the
input the modified CKY algorithm must process synthetic rules even at
positions in the chart that cross a boundary.

Thus in the modified algorithm, cells which do not cross an
independent boundary are processed normally, and in cells which do
cross a boundary the algorithm will avoid looping over complete
binary rules. While boundary-crossing cells depend on non-crossing
cells, the reverse is not the case; thus the non-crossing cells can
all be processed before the crossing cells, or the cells can be looped
over in the regular order, with a check inside the loop.

\subsection{How much work can we expect to save?}
\label{sec-3-1}

For our purposes, we can consider the amount of work done in the CKY algorithm
to be the number of binary edges visited in the inner loop. For each cell
the algorithm iterates over the binary rules in the grammar, calculating the
probability of the left-hand-side at each split point. The number of these
binary edges is

\begin{equation*}
|G|\left[\frac{n^3}{6} - \frac{n}{6}\right]
\end{equation*}

The amount of work saved depends on the number and locations of the
independent span boundaries, as well as the proportion of
non-synthetic edges in the grammar, denoted $\frac{|G_r|}{|G|}$. We
can consider two idealized scenarios: one boundary at $\frac{n}{K}$,
and $K-1$ boundaries at $\frac{n}{K}, \frac{2n}{K}, \dots,
\frac{(K-1)n}{K}$, where $K$ is an integer $1 < K < n$.

@@

\begin{equation*}
\frac{|G_r|}{|G|} \left[ \frac{3}{K} - \frac{3}{K^2} \right]
\end{equation*}

as $n$ grows. This limit converges quickly for $n \ge 10$. If we
approximate $|G_r|/|G|$ as 0.5 (for the grammar used by the parser in
Section \ref{sec:parser}, it is $\approx .54$), then for
$K=2,3,4,\dots$, the values are $\frac{3}{8}, \frac{1}{3},
\frac{3}{32}, \dots$. Intuitively, for one boundary, the best location
is exactly in the center of the sentence, and the upper limit on how
much work is saved is about 37\%.

@@

\begin{equation*}
\frac{|G_r|}{|G|}\frac{K^2 - 1}{K^2}
\end{equation*}

The values for $K=2,3,4,\dots$ are $\frac{3}{8}, \frac{4}{9},
\frac{15}{32}, \dots$. Clearly, the smaller pieces a sentence can be
divided into the less work the parser will do; however, realistically
most sentences will not have a large number of independent spans, and
they will not be equal in length. We might take $K=3$ as best-case
estimate, giving us about 44\%. Thus we can guess that a parser will be
able to save around 35-45\% of the work it does in the CKY algorithm
loop (not counting unary rules) by using independence constraints.

@@

\begin{tabular}{lrrrrrrrrrr}
Features & \#feats & Acc & Prec & Rec & F$_{\text{1}}$ & F$_{\text{0.5}}$ & TP & FP & FN & TN\\
\hline
p & 37001 & 93.71 & 80.73 & 70.49 & 75.27 & 78.45 & 3679 & 878 & 1540 & 32320\\
P$_{\text{0}}$ & 33167 & 87.16 & 51.69 & 83.98 & 63.99 & 55.99 & 4383 & 4097 & 836 & 29101\\
\hline
p,P$_{\text{0}}$ & 70168 & 95.21 & 87.38 & 75.65 & 81.09 & 84.75 & 3948 & 570 & 1271 & 32628\\
p,P$_{\text{1}}$ & 37055 & 94.81 & 78.38 & \textbf{85.38} & 81.73 & 79.69 & 4456 & 1229 & 763 & 31969\\
p,P$_{\text{2}}$ & 39336 & 95.34 & 84.25 & 80.76 & 82.47 & 83.53 & 4215 & 788 & 1004 & 32410\\
p,P$_{\text{3}}$ & 46861 & 95.04 & 89.47 & 71.95 & 79.76 & 85.31 & 3755 & 442 & 1464 & 32756\\
\hline
p,P$_{\text{0}}$,P$_{\text{1}}$ & 70222 & \textbf{95.48} & 88.95 & 76.16 & 82.06 & \textbf{86.06} & 3975 & 494 & 1244 & 32704\\
p,P$_{\text{0}}$,P$_{\text{2}}$ & 72503 & 95.09 & 88.28 & 73.60 & 80.27 & 84.89 & 3841 & 510 & 1378 & 32688\\
p,P$_{\text{0}}$,P$_{\text{3}}$ & 80028 & 94.84 & 88.81 & 70.99 & 78.91 & 84.56 & 3705 & 467 & 1514 & 32731\\
\hline
p,P$_{\text{1}}$,P$_{\text{2}}$ & 39390 & 95.27 & 80.99 & 85.21 & \textbf{83.04} & 81.80 & 4447 & 1044 & 772 & 32154\\
p,P$_{\text{1}}$,P$_{\text{3}}$ & 41553 & \textbf{95.44} & 89.05 & 75.74 & 81.86 & \textbf{86.03} & 3953 & 486 & 1266 & 32712\\
\hline
p,P$_{\text{0}}$,P$_{\text{1}}$,P$_{\text{2}}$,P$_{\text{3}}$ & 82417 & 95.35 & 86.89 & 77.49 & 81.92 & 84.83 & 4044 & 610 & 1175 & 32588\\
\end{tabular}

@@

There are several things worth noting in these results. First, neither
local nor global features are sufficient alone; it appears that local
features promote precision, while global features promote recall.
Second, examining the cases where global features are limited to a
single POS level, it is apparent that each POS level has a different
effect on precision and recall, thus confirming that the classifier is
able to extract different signals from the different POS levels, as
intended. Finally, combining all POS levels together actually reduces
accuracy, likely due to overfitting (although see the discussion of
the kernel classifier).

@@

\subsection{Polynomial Kernel}
\label{sec-4-5}

@@

\begin{itemize}
\item after reading in the grammar, index the synthetic binary rules
\item read in the file containing the boundaries output by the classifier
from the previous section
\item for each CKY cell, if the cell spans a boundary then loop over just
the synthetic binary rules
\item if at the end of the CKY loop a parse was not successful, then loop
again over just the cells which span a boundary and process all of
the binary rules
\item output the total number of times entering the inner loop as well as the
number of times the parser failed
\end{itemize}

\subsection{Experimental Setup}
\label{sec-5-1}

We use the modified Stanford Parser described above, with a grammar
extracted from the WSJ sections 02-21, and evaluate its performance on
section 22 using output from the clasifier as constraints. We vary the
threshold on the probability output by the classifier, and further
experiment with restricting the constraints to sentences above a
certain length. Finally, to compare with previous results we run the
classifier and parser on section 23 in a single configuration.

All experiments were run on a DELL Precision 690, with 8 cores and 32G
of RAM. Unless otherwise noted multiple processes were run in
parallel, and times reported were not averaged over multiple runs.
Since we saw significant variation of up to 10\%, the times should be
taken with a grain of salt. The computation done in the CKY algorithm
is measured in the number of binary edges visited in the inner loop. A
binary edge is a tuple of a span (begin \& end), a binary rule $A →
BC$, and a split point (the position where $B$ and $C$ meet).

\subsection{Results}
\label{sec-5-2}

@@

\begin{tabular}{rlrllr}
SentLen & Constraints & Time (s) & \# Edges & F$_{\text{1}}$ & Parse Failures\\
\hline
0 & default & 1283 & 1.08\texttimes{}10$^{\text{10}}$ (62\%) & 83.71 (-2.14) & 15\\
0 & precision & 1143 & 1.13\texttimes{}10$^{\text{10}}$ (65\%) & 84.05 (-1.80) & 7\\
0 & max precision & 1384 & 1.42\texttimes{}10$^{\text{10}}$ (81\%) & 85.55 (-0.30) & 2\\
0 & recall & 1024 & 7.80\texttimes{}10$^{\text{09}}$ (45\%) & 78.74 (-7.11) & 136\\
20 & default & 1126 & 1.12\texttimes{}10$^{\text{10}}$ (64\%) & 84.17 (-1.68) & 9\\
20 & precision & 1313 & 1.16\texttimes{}10$^{\text{10}}$ (66\%) & 84.43 (-1.42) & 4\\
20 & max precision & 1338 & 1.44\texttimes{}10$^{\text{10}}$ (82\%) & 85.59 (-0.26) & 2\\
20 & recall & 1121 & 8.24\texttimes{}10$^{\text{09}}$ (47\%) & 80.38 (-5.47) & 103\\
30 & default & 1312 & 1.28\texttimes{}10$^{\text{10}}$ (73\%) & 84.82 (-1.03) & 3\\
30 & precision & 1279 & 1.31\texttimes{}10$^{\text{10}}$ (75\%) & 85.01 (-0.84) & 1\\
30 & max precision & 1485 & 1.53\texttimes{}10$^{\text{10}}$ (87\%) & 85.63 (-0.22) & 1\\
30 & recall & 1140 & 1.02\texttimes{}10$^{\text{10}}$ (58\%) & 82.79 (-3.06) & 57\\
40 & default & 1476 & 1.51\texttimes{}10$^{\text{10}}$ (86\%) & 85.56 (-0.29) & 1\\
40 & precision & 1390 & 1.52\texttimes{}10$^{\text{10}}$ (87\%) & 85.59 (-0.26) & 0\\
40 & max precision & 1513 & 1.65\texttimes{}10$^{\text{10}}$ (94\%) & 85.75 (-0.10) & 0\\
40 & recall & 1403 & 1.33\texttimes{}10$^{\text{10}}$ (76\%) & 84.65 (-1.20) & 14\\
\hline
∞ & baseline & 1558 & 1.75\texttimes{}10$^{\text{10}}$ (100\%) & 85.85 & 0\\
\end{tabular}

%}
\caption{Independence constraints reduce the work done by the CKY algorithm, trading off accuracy.}
\label{tbl:parse-results-linear}
\end{table*}

The results of running the parser on section 22 using the linear
classifier from Section \ref{sec:linear-classifier} are shown in
Table \ref{tbl:parse-results-linear}. The table shows the total time
taken, the total times entering the inner loop, the F$_{\text{1}}$ and difference
from the baseline, and the number of times the parse failed using the
constraints. The baseline consisted of the same parser with the
sentence length threshold set to 1000. The time includes the time
spent reading in the constraints but not the time taken by the
classifier.

The parser with the independence constraints saves 35-38\%
of the computation inside the CKY loop over the baseline,
corresponding to about 20\% reduction in total time, at the cost of a
2-point drop in F-score. After increasing recall by making negative
instances for which the classifier assigned a low probability positive,
the parser reduced the work done inside the loop to less than half the
baseline, but accuracy also plummeted by 7 points.

@@

The results of running the parser on section 22 with the polynomial
classifier output are shown in Table \ref{tbl:parse-results-poly}.
With the more accurate classifier, the parser is able to reduce the
necessary computation even further, by about 45\%, while losing less
accuracy. With a high-precision threshold, the computation of the CKY
algorithm is reduced to less than 60\% of the baseline, while losing
less than half a point F$_{\text{1}}$ score.

\begin{tabular}{rlrllr}
SentLen & Constraints & Time (s) & \# Edges & F$_{\text{1}}$ & Parse Failures\\
\hline
0 & default & 1106 & 9.74\texttimes{}10$^{\text{09}}$ (56\%) & 84.85 (-1.00) & 6\\
0 & precision & 1118 & 9.84\texttimes{}10$^{\text{09}}$ (56\%) & 85.12 (-0.73) & 4\\
0 & max precision & 1137 & 1.02\texttimes{}10$^{\text{10}}$ (58\%) & 85.42 (-0.43) & 2\\
0 & recall & 1050 & 9.25\texttimes{}10$^{\text{09}}$ (53\%) & 84.05 (-1.80) & 33\\
20 & default & 1070 & 1.02\texttimes{}10$^{\text{10}}$ (58\%) & 85.08 (-0.77) & 5\\
20 & precision & 1172 & 1.03\texttimes{}10$^{\text{10}}$ (59\%) & 85.25 (-0.60) & 3\\
20 & max precision & 1092 & 1.06\texttimes{}10$^{\text{10}}$ (61\%) & 85.41 (-0.44) & 2\\
20 & recall & 1088 & 9.68\texttimes{}10$^{\text{09}}$ (55\%) & 84.75 (-1.10) & 7\\
30 & default & 1222 & 1.20\texttimes{}10$^{\text{10}}$ (69\%) & 85.57 (-0.28) & 1\\
30 & precision & 1267 & 1.20\texttimes{}10$^{\text{10}}$ (69\%) & 85.62 (-0.23) & 1\\
30 & max precision & 1238 & 1.23\texttimes{}10$^{\text{10}}$ (70\%) & 85.65 (-0.20) & 1\\
30 & recall & 1238 & 1.16\texttimes{}10$^{\text{10}}$ (66\%) & 85.44 (-0.41) & 2\\
40 & default & 1465 & 1.49\texttimes{}10$^{\text{10}}$ (85\%) & 85.72 (-0.13) & 0\\
40 & precision & 1353 & 1.49\texttimes{}10$^{\text{10}}$ (85\%) & 85.75 (-0.10) & 0\\
40 & max precision & 1570 & 1.50\texttimes{}10$^{\text{10}}$ (86\%) & 85.78 (-0.07) & 0\\
40 & recall & 1489 & 1.47\texttimes{}10$^{\text{10}}$ (84\%) & 85.69 (-0.16) & 1\\
\hline
∞ & baseline & 1470 & 1.75\texttimes{}10$^{\text{10}}$ (100\%) & 85.85 & 0\\
\end{tabular}

%}
\caption{The classifier using the polynomial kernel is much more accurate, leading to smaller loss in accuracy of the parser.}
\label{tbl:parse-results-poly}
\end{table*}

\subsection{WSJ Section 23}
\label{sec-5-4}

@@

\begin{center}
\begin{tabular}{lrlrr}
parser & time (s) &  & F$_{\text{1}}$ & \\
\hline
baseline & 1538 &  & 85.54 & 0.00\\
linear & 1106 & \texttimes{}1.39 & 83.55 & -1.99\\
poly & 1040 & \texttimes{}1.48 & 84.57 & -0.97\\
\end{tabular}
\end{center}

@@

Like these approaches, our method uses a classifier to avoid doing
work in certain chart cells. While not completely orthogonal, we
believe our independence constraints are complementary. A single
decision by our classifier closes a large swath of cells based on the
global structure, while their methods make local decision using local
information.

>>>>

% Created 2015-05-16 Sat 13:06
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage{acl2015}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{forest}
\usepackage[linesnumbered]{algorithm2e}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\BigO}[1]{\ensuremath{\operatorname{O}\bigl(#1\bigr)}}
\date{\today}
\title{CKY Parsing With Independence Constraints}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 24.4.1 (Org mode 8.2.10)}}
\begin{document}

@@

The CKY algorithm is an \BigO{|G|n^3} dynamic programming
algorithm for finding all of the possible derivations of a sentence in
a context-free language. Its complexity depends on both the sentence
length $n$ and the size of the grammar $|G|$. Methods for improving
parse accuracy typically increase the size of the grammar 
\cite{Klein2003,Petrov2007} or even the exponent of $n$ \cite{Eisner1999}. 
More powerful “deep” grammar formalisms multiply the computational
complexity even more \cite{Bangalore1999}.

A common technique for speeding up such parsers is coarse-to-fine
parsing, where input is first parsed using a much simpler (and thus
smaller) grammar, and the content of the chart is then used to
constrain the search over the final grammar
\cite{Torisawa2000,Charniak2005,Petrov2007}. Even with a much smaller
grammar, the CKY algorithm may be expensive---Roark et al. (2012)
report that the initial CKY step in the Berkeley Parser takes half of
the total parse time.

Other techniques can be used to prune cells in the chart. Roark et al.
(2012) use a finite-state model to label words that do/don't begin/end
spans, and skip cells that don't satisfy the labels. Bodenstab et al.
(2011) directly apply a classifier to each cell to decide how many
spans to keep. Both approaches reduce the work done by the parser
while preserving accuracy.

We propose a novel type of top-down constraint for a CFG parser that
we call independence constraints, described in Section 2. In Section 3
we show how the CKY algorithm can be easily modified to accommodate
these constraints, and in Section 4 we describe a classifier which can
provide the constraints to a parser. We integrate the constraints into
the Stanford Parser CKY implementation and show the results in Section 5.

@@

\begin{figure}
\centering
\begin{forest}
  [S
   [NP [DT [ $_0$ This $_1$]]]
   [VP
    [VB [is $_2$]]
    [NP [DT [an $_3$]]
        [NN [example $_4$]]]]
   [{.} [{.} $_5$]]
  ]
\end{forest}
\caption{In this tree `This' and `is' are independent, while `is' and `an' are not.}
\label{fig:independence}
\end{figure}

@@

An example is shown in Figure \ref{fig:independence}. Here, `This' and
`is' are independent, as are `example' and `.'. The independent spans
are (`This'), (`is', `an', `example'), and (`.'), with boundaries 1
and 4. The independent spans and independent span boundaries can be
derived straightforwardly from the definition of independent words:
the locations between consecutive words which are independent are the
independent span boundaries, and the independent spans are simply the
spans in between consecutive boundaries.

@@

The first complication is that if we assume that the independent boundaries
will be identified automatically, we must allow for errors. If a
location which is not an independent span boundary is given as one,
the parser will make an error it would not have otherwise. On the
other hand, if a location which is an independent span boundary is not
marked as such, the parser may account for this at the cost of not
achieving the minimum computation possible. By allowing for this
second type of error, the algorithm is made more robust, and allows
the independent boundary identification step to prioritize precision
over recall to lessen negative impact on the parser's accuracy.

The second issue is caused by the binarization of the context-free grammar used
in the CKY algorithm. Because the CKY algorithm requires a binary grammar, any
rules in the original grammar that have more than two symbols on the right-hand
side must be converted into a sequence of binary rules. The extra rules created
in this process are called incomplete\footnote{E.g., if a rule $A \rightarrow B C D$ becomes $@_{BC} \rightarrow B C$ and
$A \rightarrow @_{BC} D$, then the former is \emph{incomplete} and the latter is \emph{complete}.} rules. The topmost span in
particular will usually need to be constructed in several steps, applying
multiple incomplete rules before creating a complete span. If the grammar rules
are always binarized from the left (right), then only cells on the left (right)
edge of the chart can affect the top span; however, the grammar used in our
parser is binarized ``head-outward'' \cite{Klein2003}, which means that
potentially any cell in the chart can be used to create the top span.

The combination of these two issues means that in order to correctly parse a
sentence when an independent span boundary is missing from the input the
modified CKY algorithm must process incomplete rules even at positions in the
chart that cross a boundary. Thus in the modified algorithm, cells which do not
cross an independent boundary are processed normally, and in cells which do
cross a boundary the algorithm will avoid looping over complete binary rules.\footnote{While boundary-crossing cells depend on non-crossing
cells, the reverse is not the case; thus the non-crossing cells can
all be processed before the crossing cells, or the cells can be looped
over in the regular order, with a check inside the loop. While this may have
implications for e.g. parallelization, we do not explore this idea further here.}
Figure \ref{fig:chart} shows an example CKY chart where boundary-crossing cells are
colored gray.

\begin{figure}[t]
  \centering
  \includegraphics[width=6cm]{chart-constraints}
  \caption{Example of a CKY chart with independence constraints. In the gray cells the modified algorithm will only loop over incomplete rules. \label{fig:chart}}
\end{figure}

\subsection{How much work can we expect to save?}
\label{sec-3-1}
\label{sec:comp-saved}

\begin{algorithm}[t]
  \caption{The CKY algorithm. $T_{i,j}$ is the cell corresponding to words $w_i \dots w_{j-1}$.\label{alg:cky}}
  \DontPrintSemicolon
  \For {$1 \le i \le n$}{
    $T_{i,i+1} \gets \{A|A\rightarrow a \in G \wedge w_i = a\}$
  }
  \For {$2 \le j \le n$}{
    \For {$1 \le i \le n-j+1$}{
      \For {$i < k < i+j$}{
        $T_{i,i+j} \gets \{A|A\rightarrow BC \in G \wedge B \in T_{i,k} \wedge C \in T_{k,i+j} \}$\;
      }
    }
  }
\end{algorithm}

The core of the CKY algorithm is shown in Algorithm \ref{alg:cky}. For our
purposes, we can consider the amount of work done in the CKY algorithm to be the
number of binary edges visited in the inner loop (lines 4-10). For each cell the
algorithm iterates over the binary rules in the grammar, calculating the
probability of the left-hand-side at each split point. The number of these
binary edges is

\begin{equation}
|G|\left[\frac{n^3}{6} - \frac{n}{6}\right]
\end{equation}

The amount of work saved depends on the number and locations of the
independent span boundaries, as well as the proportion of
complete rules in the grammar, denoted $\frac{|G_{comp}|}{|G|}$. We
can consider two idealized scenarios: a) one boundary at $\frac{n}{K}$,
and b) $K-1$ boundaries at $\frac{n}{K}, \frac{2n}{K}, \dots,
\frac{(K-1)n}{K}$, where $K$ is an integer $1 < K < n$.

@@

\begin{equation}
\frac{|G_{comp}|}{|G|} \left[ \frac{3}{K} - \frac{3}{K^2} \right]
\end{equation}

as $n$ grows. This limit converges quickly for $n \ge 10$. If we
approximate $|G_{comp}|/|G|$ as 0.5 (for the grammar used by the parser in
Section \ref{sec:parser}, it is $\approx .54$), then for
$K=2,3,4,\dots$, the values are $\frac{3}{8}, \frac{1}{3}, \frac{3}{32}, \dots$
Intuitively, for one boundary, the best location
is exactly in the center of the sentence, and the upper limit on how
much work is saved is about 37\%.

@@

\begin{equation}
\frac{|G_{comp}|}{|G|}\frac{K^2 - 1}{K^2}
\end{equation}

The values for $K=2,3,4,\dots$ are $\frac{3}{8}, \frac{4}{9}, \frac{15}{32}, \dots$
Clearly, the smaller pieces a sentence can be
divided into the less work the parser will do; however, realistically
most sentences will not have a large number of independent spans, and
they will not be equal in length. We might take $K=3$ as best-case
estimate, giving us about 44\%. Thus we can guess that a parser will be
able to save around 35-45\% of the work it does in the CKY algorithm
loop by using independence constraints.

The derivations of Equations 1-3 are shown in the appendix.

@@

\begin{tabular}{lrrrrrrrrrr}
Features & \#feats & Acc & Prec & Rec & F$_{\text{1}}$ & F$_{\text{0.5}}$ & TP & FP & FN & TN\\
\hline
p & 37001 & 93.71 & 80.73 & 70.49 & 75.27 & 78.45 & 3679 & 878 & 1540 & 32320\\
P$_{\text{0}}$ & 33167 & 87.16 & 51.69 & 83.98 & 63.99 & 55.99 & 4383 & 4097 & 836 & 29101\\
\hline
p,P$_{\text{0}}$ & 70168 & 95.21 & 87.38 & 75.65 & 81.09 & 84.75 & 3948 & 570 & 1271 & 32628\\
p,P$_{\text{1}}$ & 37055 & 94.81 & 78.38 & 85.38 & 81.73 & 79.69 & 4456 & 1229 & 763 & 31969\\
p,P$_{\text{2}}$ & 39336 & 95.34 & 84.25 & 80.76 & 82.47 & 83.53 & 4215 & 788 & 1004 & 32410\\
p,P$_{\text{3}}$ & 46861 & 95.04 & 89.47 & 71.95 & 79.76 & 85.31 & 3755 & 442 & 1464 & 32756\\
\hline
p,P$_{\text{0}}$,P$_{\text{1}}$ & 70222 & 95.48 & 88.95 & 76.16 & 82.06 & 86.06 & 3975 & 494 & 1244 & 32704\\
p,P$_{\text{0}}$,P$_{\text{2}}$ & 72503 & 95.09 & 88.28 & 73.60 & 80.27 & 84.89 & 3841 & 510 & 1378 & 32688\\
p,P$_{\text{0}}$,P$_{\text{3}}$ & 80028 & 94.84 & 88.81 & 70.99 & 78.91 & 84.56 & 3705 & 467 & 1514 & 32731\\
\hline
p,P$_{\text{1}}$,P$_{\text{2}}$ & 39390 & 95.27 & 80.99 & 85.21 & 83.04 & 81.80 & 4447 & 1044 & 772 & 32154\\
p,P$_{\text{1}}$,P$_{\text{3}}$ & 41553 & 95.44 & 89.05 & 75.74 & 81.86 & 86.03 & 3953 & 486 & 1266 & 32712\\
\hline
p,P$_{\text{0}}$,P$_{\text{1}}$,P$_{\text{2}}$,P$_{\text{3}}$ & 82417 & 95.35 & 86.89 & 77.49 & 81.92 & 84.83 & 4044 & 610 & 1175 & 32588\\
\end{tabular}

@@

There are several things worth noting in these results. First, neither local nor
global features are sufficient alone; it appears that local features promote
precision, while global features promote recall. Second, examining the cases
where global features are limited to a single POS level, it is apparent that
each POS level has a different effect on precision and recall, thus confirming
that the classifier is able to extract different signals from the different POS
levels, as intended. Finally, combining all POS levels together actually reduces
accuracy, possibly because the features are highly correlated (although see the
discussion of the kernel classifier).

@@

\subsection{Efficiency of the Classifier}
\label{sec-4-5}

The efficiency of the classifier is as important as the accuracy---it doesn't
matter how much time is saved during parsing if it takes even longer to run the
classifier. \texttt{opal} takes less than half a second to run on the instances from
section 22; however, the instances are created by a Python script, which is not
very optimized. This script takes about 100 seconds to run on the machine
described in Section \ref{sec:setup}. While this time is already less than the
time saved in the parser (see Section \ref{sec:parse-results}), it could be
significantly reduced by reimplementing in Java or even C++. Thus the potential
gains offered by this approach are not just theoretical.

\subsection{Polynomial Kernel}
\label{sec-4-6}

@@

\begin{itemize}
\item after reading in the grammar, index the incomplete binary rules
\item read in the file containing the boundaries output by the classifier
from the previous section
\item for each CKY cell, if the cell spans a boundary then loop over just
the incomplete binary rules
\item if at the end of the CKY loop a parse was not successful, then loop
again over just the cells which span a boundary and process all of
the binary rules
\item output the total number of times entering the inner loop as well as the
number of times the parser failed
\end{itemize}

@@

\begin{tabular}{lllllr}
Parser & Time (s) & Speedup & \# Binary Edges & F$_{\text{1}}$ & Parse Failures\\
\hline
baseline & 1558 & - & 1.75\texttimes{}10$^{\text{10}}$ (100\%) & 85.85 & 0\\
linear & 1283 (+100) & 1.21\texttimes{} (1.12\texttimes{}) & 1.08\texttimes{}10$^{\text{10}}$ (62\%) & 83.71 (-2.14) & 15\\
poly & 1106 (+2h) & 1.41\texttimes{} (.19\texttimes{}) & 9.74\texttimes{}10$^{\text{09}}$ (56\%) & 84.85 (-1.00) & 6\\
oracle & 1016 & 1.53\texttimes{} & 8.47\texttimes{}10$^{\text{09}}$ (48\%) & 86.71 (+0.86) & 4\\
\end{tabular}

%}
\caption{Results of parsing with independence constraints. Results for both linear and polynomial classifiers are shown, as well as
for the gold independent span boundaries. The times in parentheses are the classifier run times.}
\label{tbl:parse-results}
\end{table*}

\subsection{Experimental Setup}
\label{sec-5-1}
\label{sec:setup}

We used the modified Stanford Parser described above, with an unlexicalized
grammar\footnote{The grammar was extracted using the Stanford Parser with command-line options \texttt{-acl03pcfg -noRebinarization -compactGrammar 1}} extracted from the WSJ sections 02-21, and evaluated its performance
on section 22 using output from the clasifier as constraints. For the baseline,
the parser was given null constraints.

All experiments were run on a DELL Precision 690, with 8 cores and 32G
of RAM. Unless otherwise noted multiple processes were run in
parallel, and times reported were not averaged over multiple runs.
Since we saw significant variation of up to 10\%, the times should be
taken with a grain of salt. The computation done in the CKY algorithm
is measured in the number of binary edges visited in the inner loop. A
binary edge is a tuple of a span (begin \& end), a binary rule $A \rightarrow BC$,
and a split point (the position where $B$ and $C$ meet).

\subsection{Results}
\label{sec-5-2}
\label{sec:parse-results}

The results of running the parser on section 22 using the linear classifier from
Section \ref{sec:linear-classifier} are shown in Table
\ref{tbl:parse-results}. The table shows the total time taken, the total
times entering the inner loop, the F$_{\text{1}}$ and difference from the baseline, and the
number of times the parse failed using the constraints. The parser with
independence constraints saves 38\% of the computation inside the CKY loop over
the baseline, corresponding to about 20\% reduction in total parse time (12\% if
the running time of the classifier is included), at the cost of a 2-point drop
in F-score.

@@

The results of running the parser on section 22 with the polynomial classifier
output are shown with the previous results in Table \ref{tbl:parse-results}.
With the more accurate classifier, the parser is able to reduce the necessary
computation even further, by 44\%, while losing less accuracy.

\subsection{Gold Independent Span Boundaries}
\label{sec-5-4}

For another comparison, we tested the parser using the gold independent span
boundaries. The results for section 22 are shown in Table
\ref{tbl:parse-results}. The number of binary edges visited is cut in half, and
parse accuracy is improved by almost 1 point. It is interesting to note that the
parser was unable to parse 4 sentences with the gold constraints (the grammar
only allowed a parse that violated the gold boundaries).

\subsection{WSJ Section 23}
\label{sec-5-5}

@@

\begin{center}
\begin{tabular}{lrlrr}
Parser & Time (s) &  & F$_{\text{1}}$ & \\
\hline
baseline & 1538 &  & 85.54 & 0.00\\
linear & 1106 & 1.39\texttimes{} & 83.55 & -1.99\\
(w/ class.) & (1206) & (1.28\texttimes{}) &  & \\
poly & 1040 & 1.48\texttimes{} & 84.57 & -0.97\\
\end{tabular}
\end{center}

As a point of comparison, Roark et al. (2012) reported speedups of 1.6-2x with
no loss of accuracy. These results are not directly comparable due to
differences in parser (their parsers use beam search variants of CYK and
coarse-to-fine pruning) and grammar (they used the Berkeley latent variable
grammar and a lexicalized grammar).

@@

Like these approaches, our method uses a classifier to avoid doing
work in certain chart cells. While not completely orthogonal, we
believe our independence constraints are complementary. A single
decision by our classifier closes a large swath of cells based on the
global structure, while their methods make local decision using local
information. The high accuracy of their classifiers shows the necessity
of improving our model.

\cite{Yarmohammadi2014} proposes a concept of `hedge' parsing, where only spans
below a certain length are allowed, and show how this reduces the computation
done by the CKY algorithm. Their system does not create spans of length larger
than the threshold and thus doesn't follow the original treebank annotation,
while our approach is able to return the original gold parse tree, provided that
the classifier does not output a false positive. Their approach of segmenting a
sentence before parsing is essentially the same as ours, but they segment based
on a maximum span length and their classifier is based on a finite-state
sequence model.
